### 2020-05-31 writeup/transfer.tex
sec 32

\begin{proposition}[Task model distance]\label{prop_dist_transition}
	In the isotropic model, suppose that $\rho_1 > 1$ and $\sigma_1 = \sigma_2 = \sigma$.
	Then we have that
	%Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between $\Psi(\beta_1, \beta_2)$ and $\Phi(\rho_1, \rho_2)$:
	\squishlist
		\item If $\Psi(\beta_1, \beta_2) < \gamma_{-} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Psi(\beta_1, \beta_2) \ge \gamma_{+} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\squishend
	Here $\gamma_{-} = (1-\oo(1))\cdot(1 - \rho_1^{-1/2})^4$ and $\gamma_{+} = (1+\oo(1))\cdot(1 + \rho_1^{-1/2})^4$.
	Concretely, if $\rho_1 > 40$, then $\gamma_{-} \in (1,2)$ and $\gamma_{+} \in (1/2, 1)$.
\end{proposition}

@@

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Task model}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Data size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Three takeaways of our theory in Section \ref{sec_insight}.
	(a) As task model distance increases, we observe a transition from positive to negative transfer  (Proposition \ref{prop_dist_transition}).
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer (Proposition \ref{prop_data_size}).
	(c) As covariate shift gets more severe (measured by increasing $\lambda$), MTL performance gets worse (Proposition \ref{prop_covariate}).}
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

\subsection{Data Size}
In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the data sizes are equal for all tasks \cite{B00,M06,MPR16}.
More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, for many applications of multi-task learning, the data sources are usually heterogeneous.
For such settings, we have observed that adding more labeled data from one task does not always help.
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using the isotropic model, we consider what happens if we vary the source task data size.

\begin{proposition}[Unnecessary source data]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 150$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\rho_1 < \frac{\rho_2-2}{\gamma_+ \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 1) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\rho_1 > \frac{\rho_2-2}{\gamma_- \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 3) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} shows that as the source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized as follows.

Given several tasks, let the data efficiency ratio be the largest factor $x$ such that the total number of labeled datapoints needed for solving all tasks can be reduced by $x$ while keeping the same performance compared to STL.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(x)$ denote the estimator trained using only $x \cdot n_i$ datapoints from every task.
The data efficiency ratio is defined as
\[ \argmin_{x\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
We quantify the data efficiency ratio of the isotropic model as follows.

>>>>

\begin{proposition}[Task model distance]\label{prop_dist_transition}
	In the isotropic model, suppose that $\rho_1 > 1$ and $\sigma_1 = \sigma_2 = \sigma$.
	Then we have that
	%Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between $\Psi(\beta_1, \beta_2)$ and $\Phi(\rho_1, \rho_2)$:
	\squishlist
		\item If $\Psi(\beta_1, \beta_2) < \gamma_{-} \cdot \Phi(\rho_1, \rho_2)$, then w.h.p. over the randomness of $X_1,X_2$, $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Psi(\beta_1, \beta_2) \ge \gamma_{+} \cdot \Phi(\rho_1, \rho_2)$, then w.h.p. over the randomness of $X_1,X_2$, $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\squishend
	Here $\gamma_{-} = (1-\oo(1))\cdot(1 - \rho_1^{-1/2})^4$ and $\gamma_{+} = (1+\oo(1))\cdot(1 + \rho_1^{-1/2})^4$.
	Concretely, if $\rho_1 > 40$, then $\gamma_{-} \in (1,2)$ and $\gamma_{+} \in (1/2, 1)$.
\end{proposition}

@@

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Task model}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Data size}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Three takeaways of our theory in Section \ref{sec_insight}.
	(a) As task model distance increases, we observe a transition from positive to negative transfer  (Proposition \ref{prop_dist_transition}).
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer (Proposition \ref{prop_data_size}).
	(c) As covariate shift gets more severe (measured by increasing $\lambda$), MTL performance gets worse (Proposition \ref{prop_covariate}).}
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

\subsection{Data Size}
In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the data sizes are equal for all tasks \cite{B00,M06,MPR16}.
More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, for many applications of multi-task learning, the data sources are usually heterogeneous.
For such settings, we have observed that adding more labeled data from one task does not always help.
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using the isotropic model, we consider what happens if we vary the source task data size.
Figure \ref{fig_size} provides a simulation result for such a setting.
We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
Our result below explains this phenomenon.

\begin{proposition}[Source/task data ratio]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 150$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} < \frac{1 - 2\rho_2^{-1}}{\gamma_+ \cdot \Psi(\beta_1, \beta_2) (1 - \rho_2^{-1}) - \rho_2^{-1}}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} > \frac{1 - 2\rho_2^{-1}}{\gamma_- \cdot \Psi(\beta_1, \beta_2) (1 - 3\rho_2^{-1}) - \rho_2^{-1}}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} describes the bias-variance tradeoff in terms of the data ratio $n_1 / n_2$.
From Figure \ref{fig_size}, we see that our result is able to explain the transition from positive to negative transfer.
There are several unexplained observations near $y = 0$ caused by the errors $\gamma_-, \gamma_+$.
The proof of Proposition \ref{prop_data_size} can be found in Appendix \ref{app_proof_31}.

We use our tools to explain a key result of Taskonomy by Zamir et al.'18 \cite{ZSSGM18}, which shows that MTL can reduce the amount of labeled data needed from each task to achieve comparable performance to STL.
Formally, let the data efficiency ratio be the largest factor $x$ such that the total number of labeled datapoints needed for solving all tasks can be reduced by $x$ while keeping the same performance compared to STL.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(x)$ denote the estimator trained using only $x \cdot n_i$ datapoints from every task.
The data efficiency ratio is defined as
\[ \argmin_{x\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
